Modeling and analyzing the Corona-virus warning app with the Isabelle Infrastructure framework

### 2020-07-11 submitted_version_paper.tex
copied from paper

\institute{Middlesex University London and\\ Technische Universit\"at Berlin\\
\email{f.kammueller@mdx.ac.uk|bianca.lutz@campus.tu-berlin.de}
}
\maketitle
\begin{abstract}
We provide a model in the Isabelle Infrastructure framework of the recently published German
Corona-virus warning app. The app supports breaking infection chains by informing users
whether they have been in close contact to an infected person. The app has a decentralized
architecture that supports anonymity of users.
We provide a formal model of the existing app with the Isabelle Infrastructure framework
to show up some natural attacks in a very abstract model. We then use the security
refinement process of the Isabelle Infrastructure framework to highlight how the use of
continuously changing ephemeral ids improves the anonymity.
\end{abstract}

@@

%The CWA
%has taken a long time to develop being published only on
%16th June 2020. It
%was a quite costly project but this was mainly due to the management of
%Telekom and SAP being in the driving seat. But the app has been designed with great
The app was designed with great
attention on privacy: a distributed architecture \cite{cwa:arch} has been adopted %after a long and
%heated debate with supporters of a central architecture. The distributed architecture is
that is
based on a very clever application design whereby clients broadcast
highly anonymized so called ``Ephemeral Ids'' via Bluetooth and
store those ids of people in close proximity.
%highly anonymized so called ``Ephemeral IDs'' at physical locations via the
%Bluetooth Low Energy Beacon protocol.
%The app saves those ids of people in close proximity.
Infected persons report their infection by uploading their ids to a central server, providing all clients the means to assess exposure risk locally,
hence, stored contact data has never to be shared.
%Infected persons report their infection to a central server,
%which, in turn, provides all clients the means to reconstruct the ids
%used by reported cases over the last 14 days.
%
%When at a later date an infected person reports his infection to a central server,
%all clients are provided the means to reconstruct the ids used by this particular
%person's device over the last 14 days.
%Therefore, exposure risks can be evaluated locally and stored contact data has never to be shared.
%
%, the unique root ID is published
%and in the daily check all mobile phones connecting to the central server download
%the root IDs of infected people. Since the Ephemeral IDs can be mapped to the root ID
%all Ephemeral IDs that have been saved over the last 14 days allow users' phones to
%regularly check whether their user has been exposed to an infected person and issue
%a warning to the user and recommendation to contact health authorities.
%The warning issued by the Corona warning app entitles to
%having a Corona test done (which at the time of writing is not normally possible).

@@

The contributions of this paper are (a) formal re-engineering of the CWA
(b) providing an additional security and privacy analysis with interactive theorem proving
certification of a novel view on the system architecture including actors, locations, and policies,
(c) a formal definition of a security refinement process that allows to improve a system
based on attacks found by the attack tree analysis and (d) an application of the refinement to
improve security of the Corona-virus app specification.

In this paper, we first provide some background in Section \ref{sec:background}:
we give a brief overview of related works
and the protocol of the German app (Section \ref{sec:history}).
We then introduce the Isabelle Infrastructure framework (Section \ref{sec:isainf}).
Next, we present our model (Section \ref{sec:model}) and analysis of privacy and
attacks (Section \ref{sec:ana}). The found attack on the first abstract specification
motivates refinement. The formal definition of refinement for the Isabelle Infrastructure
framework is introduced and illustrated on the CWA (Section \ref{sec:ref})
before drawing some conclusions (Section \ref{sec:concl}).

@@

\subsubsection{Basic DP-3T architecture}
Upon installation, the app generates secret daily seeds to derive \textit{Ephemeral Ids}
(EphIDs) from them. EphIDs are generated locally with cryptographic methods and cannot be connected
to one another but only reconstructed from the secret seed they were derived from.

@@

\subsection{Isabelle Infrastructure framework}
\label{sec:isainf}
%\TODO{Adapt: this is copied from FMBC paper}
The Isabelle Infrastructure is built in the interactive generic theorem prover
Isabelle/HOL \cite{npw:02}. As a framework, it supports formalization and proof of
systems with actors and policies. It originally emerged from verification of insider 
threat scenarios but it soon became clear that the theoretical concepts, like temporal
logic combined with Kripke structures and a generic notion of state transitions were
very suitable to be combined with attack trees into a formal security engineering process
\cite{suc:16} and framework \cite{kam:19a}.
%\subsubsection{Overview}
%\label{sec:isainfover}
Figure \ref{fig:theorystruc} gives an overview of the Isabelle Infrastructure 
framework with its layers of object-logics -- each level below embeds the one
above showing the novel contribution of this paper in blue on the top. 
\begin{figure}
\vspace{-.5cm}
  \begin{center}
  \includegraphics[scale=.45]{theory_structure.pdf}
\end{center}
\vspace{-.5cm}
\caption{Generic Isabelle Infrastructure framework applied to Corona-warning app.}
\label{fig:theorystruc}
\vspace{-.5cm}
\end{figure}

@@

\subsubsection{Instantiation of Framework}
The formal model of the Corona-warning app uses the Isabelle Infrastructure framework instantiating it
by reusing its concept of {\it actors} for users and smartphones whereby locations correspond
to physical locations. The Ephemeral Ids, their sending and change is added to Infrastructures
by slightly adapting the basic state type of infrastructure graphs and accordingly the semantic rules
for the actions move, get, and put. The details of the newly adapted Infrastructure are
presented in Section \ref{sec:model}.
Technically, an Isabelle theory file \texttt{Infrastructure.thy} builds on top of the theories for Kripke 
structures and CTL (\texttt{MC.thy}), attack trees (\texttt{AT.thy}), and security refinement 
(\texttt{Refinement.thy}). Thus all these concepts can be used to specify the formal model
for the CWA, express relevant and interesting properties, and conduct
interactive proofs (with the full support of the powerful and highly automated proof support
of Isabelle).
%The CWA theory itself is an adaptation of the Infrastructure theory of the 
%Isabelle Infrastructure framework and reuses (or slightly adapts) existing concepts. 
%In the remainder of this paper, we introduce the model that we conceived for . 
All Isabelle sources are available online \cite{kam:20gitsc}.

@@

\section{Modeling and analyzing Corona-warning app}
\label{sec:model}
\subsection{Infrastructures, Policies, and Actors}
\label{sec:infra}
% Adapted from CSF paper
The Isabelle Infrastructure framework supports the representation of infrastructures 
as graphs with actors and policies attached to nodes. These infrastructures 
are the {\it states} of the Kripke structure. % for the attack trees.

The transition between states is triggered by non-parameterized
actions \texttt{get}, \texttt{move}, and \texttt{put} 
executed by actors. 
Actors are given by an abstract type \texttt{actor} and a function 
\texttt{Actor} that creates elements of that type from identities 
(of type \texttt{string} written \texttt{''s''} in Isabelle). 
Actors are contained in an infrastructure graph type \texttt{igraph}
constructed by its constructor \texttt{Lgraph}. 
\begin{ttbox}
{\bf datatype} igraph = 
         Lgraph (location \tttimes location)set 
                 location \ttfun identity set
                 identity \ttfun (string set \tttimes string set \tttimes efid)  
                 location \ttfun string \tttimes (dlm \tttimes data) set
                 location \ttfun efid set
                 actor \ttfun location \ttfun (identity \tttimes efid) set
\end{ttbox}
In the current application of the framework to the Corona-warning app case study, 
this graph contains a set of location pairs representing the %structure 
topology of the infrastructure
as a graph of nodes and a function%\footnote{ We use the common $\lambda$-abstraction, 
%e.g. \texttt{\ttlam x. True}, to define functions with parameters, here the 
%function returning True for any input $x$.} 
that assigns a set of actor identities to each node (location) in the graph.
The third component of an \texttt{igraph} assigns the credentials to each actor: 
a triple-valued function whose first range component is a set describing the credentials 
in the possession of an actor and the second component 
is a set defining the roles the actor can take on; 
most prominently the third component is the \texttt{efid} assigned to the actor. This is initially
just a natural number but will be refined to actually represent lists of ``Ephemeral'' Ids later
when refining the specification.
\begin{ttbox}
{\bf datatype} efid = Efid nat
\end{ttbox}
The fourth component of the type \texttt{igraph}
assigns security labeled data to locations, a feature not used in the
current application.
The second to last component assigns the set of efids of all currently present smart phones
to each location of the graph. The last component finally denotes the knowledge set of each
actor for each location: a set of pairs of actors and potential ids.

@@

In the CWA, the initial values for the \texttt{igraph} components use
two locations \texttt{pub} and \texttt{shop} to define the following constants (we omit the
data map component \texttt{ex\_locs}).
\begin{ttbox}
ex_loc_ass \ttequiv (\ttlam x. if x = pub then \{''Alice'', ''Bob'', ''Eve''\}  
                 else (if x = shop then \{''Charly'', ''David''\} 
                 else \{\}))
ex_creds \ttequiv (\ttlam x. if x = ''Alice'' then (\{\}, \{\}, Efid 1) else 
              (if x = ''Bob'' then  (\{\},\{\}, Efid 2) else 
              (if x = ''Charly'' then (\{\},\{\}, Efid 3) else
              (if x = ''David'' then (\{\},\{\}, Efid 4) else
              (if x = ''Eve'' then (\{\},\{\}, Efid 5)
               else (\{\},\{\},Efid 0))))))
ex_efids \ttequiv (\ttlam x. if x = pub then \{Efid 1, Efid 2, Efid 5\}
                  else (if x = shop then \{Efid 3, Efid 4\} else \{\}))
ex_knos \ttequiv (\ttlam x. (if x = Actor ''Eve'' then (\ttlam l. \{\} else (\ttlam l. \{\})))
\end{ttbox}  
These components are wrapped up into the following \texttt{igraph}.
\begin{ttbox}
ex_graph \ttequiv
      Lgraph \{(pub, shop)\} ex_loc_ass ex_creds ex_locs ex_efids ex_knos
\end{ttbox}  
Infrastructures are given by the following datatype that 
contains an infrastructure graph of type \texttt{igraph} 
and a policy given by a function that assigns local policies over a graph to
all locations of the graph. 
\begin{ttbox}
{\bf{datatype}} infrastructure = Infrastructure igraph 
                                         [igraph, location] \ttfun policy set
\end{ttbox}
There are projection functions \texttt{graphI} and \texttt{delta} when applied
to an infrastructure return the graph and the local policies, respectively.
The function \texttt{local\_policies} gives the policy for each location \texttt{x}
over an infrastructure graph \texttt{G} as a pair: the first element of this pair is 
a function specifying the actors \texttt{y} that are entitled to perform the actions 
specified in the set which is the second element of that pair.
The local policies definition for the Corona-warning app, simply permits all actions to
all actors in both locations.
\begin{ttbox}
  local_policies G \ttequiv
         (\ttlam x. if x = pub then  \{(\ttlam y. True, \{get,move,put\})\}
          else (if x = shop then \{(\ttlam y. True, \{get,move,put\})\} else \{\}))
\end{ttbox}  
For the Corona-warning app, the initial infrastructure contains the graph \texttt{ex\_graph}
with its two locations pub and shop and is then wrapped up with the local policies
into the infrastructure \texttt{Corona\_scenario} 
that represents the ``initial'' state for the Kripke structure.
\begin{ttbox}
 Corona_scenario \ttequiv Infrastructure  ex_graph local_policies
\end{ttbox}

@@

The Privacy protection goal is to avoid deanonymization. That is, an attacker should not be able to
disambiguate the set of pairs of real Ids and EphIds. This is abstractly expressed in the predicate
identifiable.
\begin{ttbox}
identifiable eid A \ttequiv is_singleton\{(Id,Eid). (Id, Eid) \ttin A \ttand Eid = eid\}
\end{ttbox}
The predicate identifiable is used to express as the global policy `Eve cannot deanonymize an Ephemeral
Id \texttt{eid} using the gathered knowledge':
%{\bf fixes} global_policy::[infrastructure, identity] \ttfun bool
%{\bf defines}  
\begin{ttbox}
  global_policy I eid \ttequiv
             \ttneg(identifiable eid 
                ((\ttbigcap (kgra(graphI I)(Actor ''Eve'')`(nodes(graphI I))))
                 - \{(x,y). x = ''Eve''\}))
\end{ttbox}

@@

For the DP-3T Ephemeral Ids \cite{dp3t:wp}, for each day $t$ a seed $SK_t$
is used to generate a list of length $n = 24 * 60 / L$, where $L$ is the duration for which
the Ephemeral Ids are posted by the smart phone
\begin{ttbox}
  EphId1 || ... || EphIdn = PRG(PRF(SKt,``broadcast key''))
\end{ttbox} 
``where PRF is a pseudo-random function (e.g., HMAC-SHA256), ``broadcast key'' is a
fixed, public string, and PRG is a pseudorandom generator (e.g. AES in counter mode)'' \cite{dp3t:wp}.

@@

Now, the refinement shows that although the Ephemeral Ids change regularly the same attack that has been
identified on the very abstract level (fixed ids) persists.
The refinement allows refining the datatype (Id $\mapsto$ EphId) but also delivers the usual trace
refinement (behaviours of the refined systems are a subset of the traces of the abstract system).
This persistence of the attack precisely shows which part of the system behaviour is responsible for the
attack. In a second refinement step using action refinement based on this insight from the (repeated)
attack, we can exclude the dangerous attack trace.
%\footnote{Implication on the actual protection goals
%  that are endangered by the attack on privacy are discussed in the larger context of the Corona-warning
%  app (see previous Section)}.

>>>>

\institute{Middlesex University London and\\ Technische Universit\"at Berlin\\
\email{f.kammueller@mdx.ac.uk|bianca.lutz@campus.tu-berlin.de}
}
\maketitle
\begin{abstract}
We provide a model in the Isabelle Infrastructure framework of the recently published German
Corona-virus warning app (CWA). The app supports breaking infection chains by informing users
whether they have been in close contact to an infected person. The app has a decentralized
architecture that supports anonymity of users.
We provide a formal model of the existing app with the Isabelle Infrastructure framework
to show up some natural attacks in a very abstract model. We then use the security
refinement process of the Isabelle Infrastructure framework to highlight how the use of
continuously changing Ephemeral Ids (EphIDs) improves the anonymity.
\end{abstract}

@@

%The CWA
%has taken a long time to develop being published only on
%16th June 2020. It
%was a quite costly project but this was mainly due to the management of
%Telekom and SAP being in the driving seat. But the app has been designed with great
The app was designed with great
attention on privacy: a distributed architecture \cite{cwa:arch} has been adopted %after a long and
%heated debate with supporters of a central architecture. The distributed architecture is
that is
based on a very clever application design whereby clients broadcast
highly anonymized identifiers (ids)
%so called ``Ephemeral Ids''
via Bluetooth and store those ids of people in close proximity.
%highly anonymized so called ``Ephemeral IDs'' at physical locations via the
%Bluetooth Low Energy Beacon protocol.
%The app saves those ids of people in close proximity.
Infected persons report their infection by uploading their ids to a central server, providing all clients the means to assess exposure risk locally,
hence, stored contact data has never to be shared.
%Infected persons report their infection to a central server,
%which, in turn, provides all clients the means to reconstruct the ids
%used by reported cases over the last 14 days.
%
%When at a later date an infected person reports his infection to a central server,
%all clients are provided the means to reconstruct the ids used by this particular
%person's device over the last 14 days.
%Therefore, exposure risks can be evaluated locally and stored contact data has never to be shared.
%
%, the unique root ID is published
%and in the daily check all mobile phones connecting to the central server download
%the root IDs of infected people. Since the Ephemeral IDs can be mapped to the root ID
%all Ephemeral IDs that have been saved over the last 14 days allow users' phones to
%regularly check whether their user has been exposed to an infected person and issue
%a warning to the user and recommendation to contact health authorities.
%The warning issued by the Corona warning app entitles to
%having a Corona test done (which at the time of writing is not normally possible).

@@

The contributions of this paper are (a) formal re-engineering of the CWA
(b) providing an additional security and privacy analysis with interactive theorem proving
certification of a novel view on the system architecture including actors, locations, and policies,
(c) a formal definition of a security refinement process that allows to improve a system
based on attacks found by the attack tree analysis and (d) an application of the refinement to
improve security of the CWA specification.

In this paper, we first provide some background in Section \ref{sec:background}:
we give a brief overview of related works
and the protocol of CWA (Section \ref{sec:history}).
We then introduce the Isabelle Infrastructure framework (Section \ref{sec:isainf}).
Next, we present our model (Section \ref{sec:model}) and analysis of privacy and
attacks (Section \ref{sec:ana}). The found attack on the first abstract specification
motivates refinement. The formal definition of refinement for the Isabelle Infrastructure
framework is introduced and illustrated on the CWA (Section \ref{sec:ref})
before drawing some conclusions (Section \ref{sec:concl}).

@@

\subsubsection{Basic DP-3T architecture}
Upon installation, the app generates secret daily seeds to derive so-called \textit{Ephemeral Ids}
(EphIDs) from them. EphIDs are generated locally with cryptographic methods and cannot be connected
to one another but only reconstructed from the secret seed they were derived from.

@@

\subsection{Isabelle Infrastructure framework}
\label{sec:isainf}
%\TODO{Adapt: this is copied from FMBC paper}
The Isabelle Infrastructure is built in the interactive generic theorem prover
Isabelle/HOL \cite{npw:02}. As a framework, it supports formalization and proof of
systems with actors and policies. It originally emerged from verification of insider 
threat scenarios but it soon became clear that the theoretical concepts, like temporal
logic combined with Kripke structures and a generic notion of state transitions were
very suitable to be combined with attack trees into a formal security engineering process
\cite{suc:16} and framework \cite{kam:19a}.
%\subsubsection{Overview}
%\label{sec:isainfover}
Figure \ref{fig:theorystruc} gives an overview of the Isabelle Infrastructure 
framework with its layers of object-logics -- each level below embeds the one
above showing the novel contribution of this paper in blue on the top. 
\begin{figure}
\vspace{-.5cm}
  \begin{center}
  \includegraphics[scale=.45]{theory_structure.pdf}
\end{center}
\vspace{-.5cm}
\caption{Generic Isabelle Infrastructure framework applied to CWA.}
\label{fig:theorystruc}
\vspace{-.5cm}
\end{figure}

@@

\subsubsection{Instantiation of Framework}
The formal model of CWA uses the Isabelle Infrastructure framework instantiating it 
by reusing its concept of {\it actors} for users and smartphones whereby locations correspond
to physical locations. The Ephemeral Ids, their sending and change is added to Infrastructures
by slightly adapting the basic state type of infrastructure graphs and accordingly the semantic rules
for the actions move, get, and put. The details of the newly adapted Infrastructure are
presented in Section \ref{sec:model}.
Technically, an Isabelle theory file \texttt{Infrastructure.thy} builds on top of the theories for Kripke 
structures and CTL (\texttt{MC.thy}), attack trees (\texttt{AT.thy}), and security refinement 
(\texttt{Refinement.thy}). Thus all these concepts can be used to specify the formal model
for the CWA, express relevant and interesting properties, and conduct
interactive proofs (with the full support of the powerful and highly automated proof support
of Isabelle).
%The CWA theory itself is an adaptation of the Infrastructure theory of the 
%Isabelle Infrastructure framework and reuses (or slightly adapts) existing concepts. 
%In the remainder of this paper, we introduce the model that we conceived for . 
All Isabelle sources are available online \cite{kam:20gitsc}.

@@

\section{Modeling and analyzing CWA}
\label{sec:model}
\subsection{Infrastructures, Policies, and Actors}
\label{sec:infra}
% Adapted from CSF paper
The Isabelle Infrastructure framework supports the representation of infrastructures 
as graphs with actors and policies attached to nodes. These infrastructures 
are the {\it states} of the Kripke structure. % for the attack trees.

The transition between states is triggered by non-parameterized
actions \texttt{get}, \texttt{move}, and \texttt{put} 
executed by actors. 
Actors are given by an abstract type \texttt{actor} and a function 
\texttt{Actor} that creates elements of that type from identities 
(of type \texttt{string} written \texttt{''s''} in Isabelle). 
Actors are contained in an infrastructure graph type \texttt{igraph}
constructed by its constructor \texttt{Lgraph}. 
\begin{ttbox}
{\bf datatype} igraph = 
         Lgraph (location \tttimes location)set 
                 location \ttfun identity set
                 identity \ttfun (string set \tttimes string set \tttimes efid)  
                 location \ttfun string \tttimes (dlm \tttimes data) set
                 location \ttfun efid set
                 actor \ttfun location \ttfun (identity \tttimes efid) set
\end{ttbox}
In the current application of the framework to the CWA case study, 
this graph contains a set of location pairs representing the %structure 
topology of the infrastructure
as a graph of nodes and a function%\footnote{ We use the common $\lambda$-abstraction, 
%e.g. \texttt{\ttlam x. True}, to define functions with parameters, here the 
%function returning True for any input $x$.} 
that assigns a set of actor identities to each node (location) in the graph.
The third component of an \texttt{igraph} assigns the credentials to each actor: 
a triple-valued function whose first range component is a set describing the credentials 
in the possession of an actor and the second component 
is a set defining the roles the actor can take on; 
most prominently the third component is the \texttt{efid} assigned to the actor. This is initially
just a natural number but will be refined to actually represent lists of Ephemeral Ids later
when refining the specification.
\begin{ttbox}
{\bf datatype} efid = Efid nat
\end{ttbox}
The fourth component of the type \texttt{igraph}
assigns security labeled data to locations, a feature not used in the
current application.
The second to last component assigns the set of efids of all currently present smart phones
to each location of the graph. The last component finally denotes the knowledge set of each
actor for each location: a set of pairs of actors and potential ids.

@@

In the CWA, the initial values for the \texttt{igraph} components use
two locations \texttt{pub} and \texttt{shop} to define the following constants (we omit the
data map component \texttt{ex\_locs}).
\begin{ttbox}
ex_loc_ass \ttequiv (\ttlam x. if x = pub then \{''Alice'', ''Bob'', ''Eve''\}  
                 else (if x = shop then \{''Charly'', ''David''\} 
                 else \{\}))
ex_creds \ttequiv (\ttlam x. if x = ''Alice'' then (\{\}, \{\}, Efid 1) else 
              (if x = ''Bob'' then  (\{\},\{\}, Efid 2) else 
              (if x = ''Charly'' then (\{\},\{\}, Efid 3) else
              (if x = ''David'' then (\{\},\{\}, Efid 4) else
              (if x = ''Eve'' then (\{\},\{\}, Efid 5)
               else (\{\},\{\},Efid 0))))))
ex_efids \ttequiv (\ttlam x. if x = pub then \{Efid 1, Efid 2, Efid 5\}
                  else (if x = shop then \{Efid 3, Efid 4\} else \{\}))
ex_knos \ttequiv (\ttlam x. (if x = Actor ''Eve'' then (\ttlam l. \{\} else (\ttlam l. \{\})))
\end{ttbox}  
These components are wrapped up into the following \texttt{igraph}.
\begin{ttbox}
ex_graph \ttequiv
      Lgraph \{(pub, shop)\} ex_loc_ass ex_creds ex_locs ex_efids ex_knos
\end{ttbox}  
Infrastructures are given by the following datatype that 
contains an infrastructure graph of type \texttt{igraph} 
and a policy given by a function that assigns local policies over a graph to
all locations of the graph. 
\begin{ttbox}
{\bf{datatype}} infrastructure = Infrastructure igraph 
                                         [igraph, location] \ttfun policy set
\end{ttbox}
There are projection functions \texttt{graphI} and \texttt{delta} when applied
to an infrastructure return the graph and the local policies, respectively.
The function \texttt{local\_policies} gives the policy for each location \texttt{x}
over an infrastructure graph \texttt{G} as a pair: the first element of this pair is 
a function specifying the actors \texttt{y} that are entitled to perform the actions 
specified in the set which is the second element of that pair.
The local policies definition for the CWA, simply permits all actions to
all actors in both locations.
\begin{ttbox}
  local_policies G \ttequiv
         (\ttlam x. if x = pub then  \{(\ttlam y. True, \{get,move,put\})\}
          else (if x = shop then \{(\ttlam y. True, \{get,move,put\})\} else \{\}))
\end{ttbox}  
For the CWA, the initial infrastructure contains the graph \texttt{ex\_graph}
with its two locations pub and shop and is then wrapped up with the local policies
into the infrastructure \texttt{Corona\_scenario} 
that represents the ``initial'' state for the Kripke structure.
\begin{ttbox}
 Corona_scenario \ttequiv Infrastructure  ex_graph local_policies
\end{ttbox}

@@

The Privacy protection goal is to avoid deanonymization. That is, an attacker should not be able to
disambiguate the set of pairs of real ids and EphIDs. This is abstractly expressed in the predicate
identifiable.
\begin{ttbox}
identifiable eid A \ttequiv is_singleton\{(Id,Eid). (Id, Eid) \ttin A \ttand Eid = eid\}
\end{ttbox}
The predicate identifiable is used to express as the global policy `Eve cannot deanonymize an Ephemeral
Id \texttt{eid} using the gathered knowledge':
%{\bf fixes} global_policy::[infrastructure, identity] \ttfun bool
%{\bf defines}  
\begin{ttbox}
  global_policy I eid \ttequiv
             \ttneg(identifiable eid 
                ((\ttbigcap (kgra(graphI I)(Actor ''Eve'')`(nodes(graphI I))))
                 - \{(x,y). x = ''Eve''\}))
\end{ttbox}

@@

For the DP-3T Ephemeral Ids \cite{dp3t:wp}, for each day $t$ a seed $SK_t$
is used to generate a list of length $n = 24 * 60 / L$, where $L$ is the duration for which
the Ephemeral Ids are posted by the smart phone
\begin{ttbox}
  EphID1 || ... || EphIDn = PRG(PRF(SKt,``broadcast key''))
\end{ttbox} 
``where PRF is a pseudo-random function (e.g., HMAC-SHA256), ``broadcast key'' is a
fixed, public string, and PRG is a pseudorandom generator (e.g. AES in counter mode)'' \cite{dp3t:wp}.

@@

Now, the refinement shows that although the Ephemeral Ids change regularly the same attack that has been
identified on the very abstract level (fixed ids) persists.
The refinement allows refining the datatype (Id $\mapsto$ EphID) but also delivers the usual trace
refinement (behaviours of the refined systems are a subset of the traces of the abstract system).
This persistence of the attack precisely shows which part of the system behaviour is responsible for the
attack. In a second refinement step using action refinement based on this insight from the (repeated)
attack, we can exclude the dangerous attack trace.
%\footnote{Implication on the actual protection goals
%  that are endangered by the attack on privacy are discussed in the larger context of the Corona-warning
%  app (see previous Section)}.
